### 2020-06-08 paper/figures/cocoa_illustration_too_much.tex

\input{cocoa_common_beginning.tex}
%\fill[pattern=flexible hatch, pattern color=red, hatch distance=1mm, hatch thickness=0.01mm] (0,0) rectangle ++(2.5,-0.5);
\draw[draw opacity=0,fill=red,fill opacity=0.3] (0,0) rectangle ++(2.5,-0.5);
%\draw[name path=A] (0,-0.5)-- ++(2.5, 0);

\draw[draw,decorate,decoration={brace,mirror}] (0.75,0) -- (0.75,-0.5) node [midway,left,align=right,font=\fontsize{3pt}{0}\selectfont] {standing\\ queue};
%\draw[densely dotted] (0.75,0) -- (0.75,-0.5);

\draw[dashed] (0,-0.5) -- ++(2.5,0);
\node[align=center,font=\fontsize{3pt}{0}\selectfont,inner sep=0mm] at (1.95,-0.25) (optimal) {optimal minimum\\ window (BDP)};
\draw[->, line width=0.05mm] (optimal) -- ++(0,-0.25);

\draw[->, line width=0.05mm] (0,-0.75) -- ++(2.5,0);
\node[align=center,font=\fontsize{3pt}{0}\selectfont,inner sep=0mm] at (2.35, -0.7) (time) {time};
\draw[draw=none, line width=0.05mm] (0,-0.8) -- ++(2.5,0);

\draw[->, line width=0.05mm, -dot3-=0.055] (0,-0.75) -- ++(0,2.1);
\node[anchor=north west, align=center,font=\fontsize{3pt}{0}\selectfont,inner sep=0mm] at (0.01, 1.3) (window) {window (packets in flight)};
\draw[draw=none, line width=0.05mm] (-0.05,-0.75) -- ++(0,2.1);

\input{cocoa_common_end.tex}