### 2020-06-10 paper/figures/cocoa_illustration_perfect.tex
Update paper draft

\input{cocoa_common_beginning.tex}

\draw[draw,decorate,decoration={brace, mirror}] (2.5,0) -- ++(0.0,1) node [midway,right,align=left,font=\fontsize{3pt}{0}\selectfont] {buffer\\ size};

\draw[dashed] (0,0) -- ++(2.5,0);
\node[align=center,font=\fontsize{3pt}{0}\selectfont,inner sep=0mm] at (1.95,-0.25) (optimal) {optimal minimum\\ window (BDP)};
\draw[->, line width=0.05mm] (optimal) -- ++(0,0.25);

\draw[->, line width=0.05mm] (0,-1) -- ++(2.5,0);
\node[align=center,font=\fontsize{3pt}{0}\selectfont,inner sep=0mm] at (2.35, -0.95) (time) {time};
\draw[draw=none, line width=0.05mm] (0,-0.8) -- ++(2.5,0);

\draw[->, line width=0.05mm] (0,-1) -- ++(0,2.35);
\node[anchor=north west, align=center,font=\fontsize{3pt}{0}\selectfont,inner sep=0mm] at (0.01, 1.3) (window) {window (packets in flight)};
\draw[draw=none, line width=0.05mm] (-0.05,-0.75) -- ++(0,2.1);

\input{cocoa_common_end.tex}